The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
More refinements on content flow
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1515 4525493e-7705-40b1-a816-d608a930855b

\section{CMS data model\label{DataModel}}
\section{Data Aggregation System\label{DAS}}
Data Aggregation System shown in Figure \ref{DAS_arch} and
consists out of several components:
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_architecture.pdf}
\caption{
DAS architecture.
}
\label{DAS_arch}
\end{figure}

\begin{itemize}
\item DAS core system which is support pluggable modules to retrieve data from
CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
etc.;
\item DAS caching system which has dual purpose: to store ``raw'' results coming
out from data-services, a raw-cache, and aggregated results produced by DAS core into
hot-cache;
\item DAS request service who collect and place into queue user requests
\item DAS mapping DB to store mapping between data-services and DAS notations and
data-representation;
\item DAS analytics DB to collect query statistics and establish pre-fetching 
strategies.
\end{itemize}
We will discuss each of those components in more details in up-coming sub-sections.
The DAS supports a pluggable approach to add a new CMS sub-system into a game by
identifying appropriate APIs, input and output parameters and keys, data-service
format and mapping between data-service, DAS and other sub-system. For instance,
the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
a file. In former case, a file name was stored as well as its relations to other physics 
objects, such as run and luminosity, while in later case the Phedex store information 
about file location. When user specify a file as their selection key, based on
information store in DAS mapping DB we were able to identify appropriate set of
data-service APIs, look-up data, re-format their output to DAS notations and
store it into DAS raw-cache. Even though we support a data retrieval on demand
basis we also store user queries into DAS analytics DB to identify
a common patterns. Based on this information we organized a pre-fetching
populators. Once data were coming to raw-cache an additional step of
aggregation was applied. If other data with identical keys were presented in
a raw-cache we merge it them together. For instance upon the following query
$find file where ... $ a file information from DBS and Phedex systems were
retrieved. Since both data-services output contained file name as a common
key, their output where merged and final result was stored into raw-cache.
This merged (aggregated) results were shipped back to client and stored into
hot-cache (memcached) to speed up a DAS web client application. Each DAS data
object contains a standard DAS header, which identify look-up time,
request url, API, parameters and method as well as response version, expiration
time and checksum of the output.

\subsection{DAS Query Language}
Upon success of DBS-QL \ref{DBS-QL} we adopted its syntax to DAS with a few minor
changes. Here we briefly recall its syntax:
\begin{equation}
action\,\,\,
key_1,\, key_2,\, ...\,\,\, where\,\,\,
\langle key\rangle\, 
\langle op\rangle\, 
\langle value\rangle \,\,\, and|or ...
\label{QL_syntax}
\end{equation}
Each DAS QL expression starts with special keyword indicating an action such as 
$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
the behavior of the query, find to look-up data, plot to plot found data, summary to
apply a summary view over provided selection keys. An optional conditions can be 
specified via $where$ keyword following by condition expression.
Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
$key$ is DAS selection key, $operator$ is conditional operator, 
$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
We also support brackets as well as $and$ and $or$ logical operators.
The DAS selection and condition keys were identified based on analysis 
of terminology used by our users. For instance, $file$ was used to either
select file name or used in condition to identify file name provided by user.
Here is a simple example of DAS QL expression:
$$
find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
$$
which represents a natural language question: 
{\it find files and runs located at T1\_CH\_CERN site}.

\subsection{DAS Mapping DB}
As been discussed above we used a DAS mapping DB to map data-service API into
DAS as well as map provide a relations between different data-services.
Upon identification of APIs which participate in DAS activity 
we store a mapping between data-service notations
into DAS mapping DB. This allows to map user input query parameters into
data-service API metrics as well as map back data-service output into DAS names.
For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
the same entity was named as $lfn$. Since both of them were referred to
logical file name, we map both notations into $file$ notation in DAS.\footnote{We
also deal with physical file names, but still used $file$ to refer to them in
DAS notation, since simple regular expressions were used to identify $file$
meaning from the provided value.}
For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
common data-format used in DAS.\footnote{We used JSON as a DAS data format}
We also store a mapping between DAS keywords and data-service API names to
establish appropriate calls to data-service upon provided DAS query. For example,
when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
DAS core library calls from hardcoded data-service API calls and make it
transparent to any participating data-services. Here is a logic of DAS core
library:
\begin{itemize}
\item parse input DAS query
\item map provided selection and condition keys into data-service APIs and
input parameters
\item invoke data-service APIs
\item parse output results by mapping output notation into DAS keys
\item store and aggregate them into DAS raw-cache
\end{itemize}

\subsection{DAS Analytics DB}

\subsection{Caching system}
The DAS caching system was divided into two different caches: raw-cache and hot-cache.
The purpose of raw-cache was to store results coming from data services for
as long as specified in data-service response header. While the purpose of hot-cache
was to keep final DAS results corresponding to provided user query. We use a populator
services (jobs) to feed raw-cache with highly requested queries. Here we will only
discuss the raw-cache organization. 
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
\caption{
DAS cache system.
}
\label{DAS_cache}
\end{figure}
As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
On one side it talks to DAS Analytics DB to trigger jobs for worker robot
who update cache with most popular queries, on another it talks to DAS
raw-cache back-end to either stored mapped results or merge them with 
existing DAS objects in a cache. We should note that we use on-demand
approach to answer user queries, but at the same time with a help of
DAS Analytics DB we force worker robots (populators) to pre-fetch
the most popular queries up-front.

@@

\subsection{REST interface}
We designed DAS interface to be fully REST complaint \cite{REST}.
All requests placed into DAS cache server were done via GET/PUT/POST/DELETE methods.
DAS clients initiate their request via GET method. An appropriate message
with either resulting data or their absense was send back to client.
Clients were able to decided upon received message how to proceed, either
request data via POST/PUT methods, DELETE data in cache or initiate another
GET request. The REST complaince help us to treat our clients, user quyering
DAS system via web browser, other CMS applications and command line tools
in identical manner. Also, the implementation of pre-fetch populator
tools become as easy and identical as any other clients.

>>>>

\section{CMS data model\label{DataModel}}
\section{Data Aggregation System\label{DAS}}
Our goal was to create scalable and pluggable system for existing data-services
within CMS computing infrastructure. We knew that all our data are delivered 
via data-services who take care of their availability. Therefore, we treat 
DAS as aggregation and cache service for existing data-services who should
take care of the following:
\begin{itemize}
\item an individual data-service can have an own set of security policies;
\item a dynamic nature of data-service APIs, due to continiously changing
detector and software environments;
\item a broad variety of data-formats;
\item distributed nature of data-services;
\item heterogeneous software environment.
\item support flexible and intuitive Query Language (QL) to easy look-up our
data;
\item pluggable interface to existing and up-coming data-service APIs;
\end{itemize}

\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_architecture.pdf}
\caption{
DAS architecture.
}
\label{DAS_arch}
\end{figure}

Based on this set of constrains and requirements we come up with the following
DAS components, see Figure \ref{DAS_arch}:
\begin{itemize}
\item DAS core system which is support pluggable modules to retrieve data from
CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
etc.;
\item DAS caching system which has dual purpose: to store ``raw'' results coming
out from data-services, a raw-cache, and aggregated results produced by DAS core into
hot-cache;
\item DAS request service to handle user requests
\item DAS mapping DB to store mapping between data-services and DAS notations and
data-representation;
\item DAS analytics DB to collect query statistics and establish pre-fetching 
strategies.
\end{itemize}
We will discuss individual components in more details in up-coming sub-sections.

\subsection{Dealing with data-services}
The DAS supports a pluggable approach to add a new CMS sub-system into a game by
identifying appropriate APIs, input and output parameters and keys, data-service
format and mapping between data-service, DAS and other sub-system. For instance,
the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
a file. In former case, a file name was stored as well as its relations to other physics 
objects, such as run and luminosity, while in later case the Phedex store information 
about file location. When user specify a file as their selection key, based on
information store in DAS mapping DB we were able to identify appropriate set of
data-service APIs, look-up data, re-format their output to DAS notations and
store it into DAS raw-cache. Even though we support a data retrieval on demand
basis we also store user queries into DAS analytics DB to identify
a common patterns. Based on this information we organized a pre-fetching
populators. Once data were coming to raw-cache an additional step of
aggregation was applied. If other data with identical keys were presented in
a raw-cache we merge it them together. For instance upon the following query
$find file where ... $ a file information from DBS and Phedex systems were
retrieved. Since both data-services output contained file name as a common
key, their output where merged and final result was stored into raw-cache.
This merged (aggregated) results were shipped back to client and stored into
hot-cache (memcached) to speed up a DAS web client application. Each DAS data
object contains a standard DAS header, which identify look-up time,
request url, API, parameters and method as well as response version, expiration
time and checksum of the output.

We designed DAS interface to be fully REST complaint \cite{REST}. It is not
only simplify development of DAS clients (CLI, web interface, populator, etc.)
but also consider DAS itself as own sub-component, i.e. place queries against
DAS content, rather then data-service one.

\subsection{Query Language\label{DAS-QL}}
Upon success of DBS-QL \ref{DBS-QL} we adopted its syntax to DAS with a few minor
changes. Here we briefly recall its syntax:
\begin{equation}
action\,\,\,
key_1,\, key_2,\, ...\,\,\, where\,\,\,
\langle key\rangle\, 
\langle op\rangle\, 
\langle value\rangle \,\,\, and|or ...
\label{QL_syntax}
\end{equation}
Each DAS QL expression starts with special keyword indicating an action such as 
$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
the behavior of the query, find to look-up data, plot to plot found data, summary to
apply a summary view over provided selection keys. An optional conditions can be 
specified via $where$ keyword following by condition expression.
Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
$key$ is DAS selection key, $operator$ is conditional operator, 
$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
We also support brackets as well as $and$ and $or$ logical operators.
The DAS selection and condition keys were identified based on analysis 
of terminology used by our users. For instance, $file$ was used to either
select file name or used in condition to identify file name provided by user.
Here is a simple example of DAS QL expression:
$$
find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
$$
which represents a natural language question: 
{\it find files and runs located at T1\_CH\_CERN site}.

\subsection{Mapping}
As been discussed above we used a DAS mapping DB to map data-service API into
DAS as well as map provide a relations between different data-services.
Upon identification of APIs which participate in DAS activity 
we store a mapping between data-service notations
into DAS mapping DB. This allows to map user input query parameters into
data-service API metrics as well as map back data-service output into DAS names.
For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
the same entity was named as $lfn$. Since both of them were referred to
logical file name, we map both notations into $file$ notation in DAS.\footnote{We
also deal with physical file names, but still used $file$ to refer to them in
DAS notation, since simple regular expressions were used to identify $file$
meaning from the provided value.}
For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
common data-format used in DAS.\footnote{We used JSON as a DAS data format}
We also store a mapping between DAS keywords and data-service API names to
establish appropriate calls to data-service upon provided DAS query. For example,
when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
DAS core library calls from hardcoded data-service API calls and make it
transparent to any participating data-services. Here is a logic of DAS core
library:
\begin{itemize}
\item parse input DAS query
\item map provided selection and condition keys into data-service APIs and
input parameters
\item invoke data-service APIs
\item parse output results by mapping output notation into DAS keys
\item store and aggregate them into DAS raw-cache
\end{itemize}

\subsection{Caching system and Analytics}
The DAS caching system was divided into two different caches: raw-cache and hot-cache.
The purpose of raw-cache was to store results coming from data services for
as long as specified in data-service response header. While the purpose of hot-cache
was to keep final DAS results corresponding to provided user query. We use a populator
services (jobs) to feed raw-cache with highly requested queries. Here we will only
discuss the raw-cache organization. 
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
\caption{
DAS cache system.
}
\label{DAS_cache}
\end{figure}
As it shown in figure \ref{DAS_cache} the DAS cache server has dual purpose.
On one side it talks to DAS Analytics DB to trigger jobs for worker robot
who update cache with most popular queries, on another it talks to DAS
raw-cache back-end to either stored mapped results or merge them with 
existing DAS objects in a cache. We should note that we use on-demand
approach to answer user queries, but at the same time with a help of
DAS Analytics DB we force worker robots (populators) to pre-fetch
the most popular queries up-front.

@@
